---
title: "Practical Machine Learning Course Project - Prediction Assignment Writeup"
author: "Ing. Raúl Mascaró"
date: "August 29, 2016"
output:
  html_document
---

##Practical Machine Learning - Prediction Assignment Writeup

For this assignment I analyzed the provided data to determine what activity an individual perform.
To do this I made use of caret and randomForest, this allowed me to generate correct answers for
each of the 20 test data cases provided in this assignment.  I made use of a seed value for 
consistent results.


```{r warning=FALSE, message=FALSE}
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(2048)
options(warn=-1)
```

The first step after loading libraries is loading the source data (for training and testing) from its url.
Values with formula errors are replaced with NA values and some columns could be casted to be numeric.

```{r warning=FALSE, message=FALSE}
train.url ="https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test.url = "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training_data <- read.csv(train.url, na.strings=c("#DIV/0!") )
evaluation_data <- read.csv(test.url, na.strings=c("#DIV/0!") )
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}
for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}
```

Columns that are not completed have been deleted. User name, timestamps and windows were also removed.  

Next, determine and display out feature set.

```{r warning=FALSE, message=FALSE}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

Model Data is now ready to be used.

```{r warning=FALSE, message=FALSE}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```
##Random Forests
We process in paralell 5 Random Forests with 150 Trees each one. 

```{r warning=FALSE, message=FALSE}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree)
}
```

Provide error reports for both training and test data (Confucion Matrices).
```{r warning=FALSE, message=FALSE}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

##Conclusions

This model have a high accuracy as we can see the confusion matrix.
Test data had an accurancy around 99% so we could expect that nearly all of the submitted test cases will be correct as finally occured.  

##TD Submission

Files for submission.

```{r warning=FALSE, message=FALSE}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```